Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-12-02 leadership_across_cultures_contexts.Rmd
Commeted out readacted sections.

## Principle components analyses

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)). We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)). Measures of physical formidability are also moderately associated with the "dominance" side of this component. The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*. Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.

```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components. PC 1 is identified as a *Prestige vs. Dominance* component and PC 2 is identified as an *In-group favoritism vs. Social Status* component."}
logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2)
```

Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance. The first leader functions component distinguishes between social functions (a very broad and generalized measure) from other more specific leader functions, which we identify as a *Social functions* component. Low values on Leader function PC 1 are strongly associated with *Social functions*. The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as an *Organization vs. Mediation* component. Low values on Leader functions PC 2 are strongly associated with the variables *Organize cooperation* and *Political appointments*, whereas high values are strongly associated with *Resolve conflict* and moderately associated with *Provide counsel/direction*. See Figure \@ref(fig:functionsPCplot).

```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components. PC 1 is identified as a *Social functions* component and PC 2 is identified as an *Organization vs. Mediation* component."}
logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2)
```

@@

## Exploring population-level variation

To explore potential population variation in the *Prestige vs. Dominance* and *In-group favoritism vs. Social Status* leader qualities components and the *Social functions* and *Mediation vs. Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures. For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:groupsex)).

@@

Functions PC 1 model

```{r funcPCmodel1, fig.height=8, fig.width=20}
m_fPC1_sub_plot + m_fPC1_group_plot
```

Functions PC 2 model

```{r funcPCmodel2, fig.height=10, fig.width=20}
m_fPC2_sub_plot + m_fPC2_group_plot
```

Qualities PC 1 model

```{r qualPCmodel1, fig.height=10, fig.width=20}
m_qPC1_sub_plot + m_qPC1_group_plot
```

Qualities PC 2 model

```{r qualPCmodel2, fig.height=10, fig.width=20}
m_qPC2_sub_plot + m_qPC2_group_plot
```

>>>>

<!-- ## Principle components analyses -->

<!-- A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)). We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)). Measures of physical formidability are also moderately associated with the "dominance" side of this component. The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*. Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status. -->

<!-- ```{r qualitiesPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader qualitiy variables on the first two components. PC 1 is identified as a *Prestige vs. Dominance* component and PC 2 is identified as an *In-group favoritism vs. Social Status* component."} -->
<!-- logisticPCA_loadings_plot(logpca_model_qualities, data = pca_data_qualities2) -->
<!-- ``` -->

<!-- Similarly, A logistic PCA on the `r length(function_vars)` coded leader function variables, with rank $k=`r dim(logpca_model_functions$U)[2]`$ and $m = `r logpca_model_functions$m`$ optimized by cross-validation, explained `r signif(logpca_model_functions$prop_deviance_expl, 3)*100`% of the deviance. The first leader functions component distinguishes between social functions (a very broad and generalized measure) from other more specific leader functions, which we identify as a *Social functions* component. Low values on Leader function PC 1 are strongly associated with *Social functions*. The Leader functions PC 2 distinguished between organizational and mediation-based functions, which we identify as an *Organization vs. Mediation* component. Low values on Leader functions PC 2 are strongly associated with the variables *Organize cooperation* and *Political appointments*, whereas high values are strongly associated with *Resolve conflict* and moderately associated with *Provide counsel/direction*. See Figure \@ref(fig:functionsPCplot). -->

<!-- ```{r functionsPCplot, fig.width=15, fig.height=10, fig.cap="Logistic PCA loadings of coded leader function variables on the first two components. PC 1 is identified as a *Social functions* component and PC 2 is identified as an *Organization vs. Mediation* component."} -->
<!-- logisticPCA_loadings_plot(logpca_model_functions, data = pca_data_functions2) -->
<!-- ``` -->

@@

<!-- ## Exploring population-level variation -->

<!-- To explore potential population variation in the *Prestige vs. Dominance* and *In-group favoritism vs. Social Status* leader qualities components and the *Social functions* and *Mediation vs. Organization* leader functions components, we fit four hierarchical linear models at the text-record level, using nested random effects of document within culture with each of these components as outcome measures. For predictor variables we relied on our researcher-coded measure of group structure and our modified version of the HRAF subsistence classification (see Figure \@ref(fig:groupsex)).  -->

@@

<!-- Functions PC 1 model -->

<!-- ```{r funcPCmodel1, fig.height=8, fig.width=20} -->
<!-- m_fPC1_sub_plot + m_fPC1_group_plot -->
<!-- ``` -->

<!-- Functions PC 2 model -->

<!-- ```{r funcPCmodel2, fig.height=10, fig.width=20} -->
<!-- m_fPC2_sub_plot + m_fPC2_group_plot -->
<!-- ``` -->

<!-- Qualities PC 1 model -->

<!-- ```{r qualPCmodel1, fig.height=10, fig.width=20} -->
<!-- m_qPC1_sub_plot + m_qPC1_group_plot -->
<!-- ``` -->

<!-- Qualities PC 2 model  -->

<!-- ```{r qualPCmodel2, fig.height=10, fig.width=20} -->
<!-- m_qPC2_sub_plot + m_qPC2_group_plot -->
<!-- ``` -->
